Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States

### 2024-11-04 paper/paper.qmd
fix appendix and type out survey

---
title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: >
  This study employs a poll-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college. By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris. Our analysis also shows a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
format: 
  pdf:
    number-sections: true
    header-includes:
      - \usepackage{pdfpages}
      - '\KOMAoption{captions}{tableheading}'
      - \usepackage[margin=1in]{geometry}

bibliography: references.bib
---

@@

This paper is organized as follows: In Section @sec-data, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables. In Section @sec-model, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions. In Section @sec-discussion, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work. [Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024. [Appendix -@sec-idealized-meth] contains a discussion about an idealized methodology that we would use to forecast the 2024 U.S. Presidential Election if we had a budget of  \$100,000.

@@

@@

# Copy of the Survey

\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}

# Model Diagnostics

The diagnostic plots for Model 3ni sgiven in [@fig-dgplotone]. We notice the following:

# Code Styling

@@

# Reproducibility

@@

# Acknowledgments

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: >
  This study employs a poll-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college. By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris. Our analysis also shows a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper is organized as follows: In Section @sec-data, we show summary statistics, plot distributions of key variables, and examine relationships between variables. In Section @sec-model, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions. In Section @sec-discussion, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work. [Appendix -@sec-data-model] contains details of the data cleaning process and model diagnostics. [Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024. [Appendix -@sec-idealized-meth] contains a discussion about an idealized methodology that we would use to forecast the 2024 U.S. Presidential Election if we had a budget of  \$100,000. [Appendix -@sec-copy] contains a copy of a survey that would be used in the idealized methodology. [Appendix -@sec-styling] contains specifications on code styling. [Appendix -@sec-reproducibility] contains notes on reproducibility and [Appendix -@acknowledgment] acknowledges important ressources consulted for the production of this paper.

@@

# Data Cleaning and Modeling {#sec-data-model}

## Model Diagnostics

The diagnostic plots for Model 3 is given in [@fig-dgplotone]. We notice the following:

@@

# Copy of the Survey {#sec-copy}

\begingroup
\small
**2024 U.S. Presidential Election Survey**

**Introduction:**
Thank you for your interest in the 2024 U.S. Presidential Election Survey. The 2024 U.S. Presidential Election is set to take place on November 5, 2024. This election will decide the next President of the United States, a position that holds significant influence over national and global policies.

Your participation in this survey is entirely voluntary, and you can choose to withdraw at any time without any consequences. The information collected will be used solely for research purposes to gain insights into voter preferences and trends as we approach this important election. All responses will remain confidential and will be reported only in aggregate form.

**Contact Information:**

- Denise Chang, University of Toronto. Email: dede.chang@mail.utoronto.ca

- Shamayla Durrin, University of Toronto. Email: shamayla.islam@mail.utoronto.ca

- Krishna Kumar, University of Toronto. Email: krishna.kumar@mail.utoronto.ca

---

**Section 1: Voting Intention**

1. Consent to Participate
By clicking the button below, I consent to participate in this survey, and I agree that my responses may be used for research purposes, including analysis and reporting on voter sentiment related to the 2024 U.S. Presidential Election. I understand that my participation is voluntary and that I can withdraw at any time without any consequences. My responses will be kept confidential and will be used only in aggregate form.
   - I agree

2. How likely are you to vote in the 2024 U.S. Presidential Election?
   - Not likely at all
   - Not very likely
   - Somewhat likely
   - Extremely likely
   - Prefer not to say

3. If the 2024 U.S. Presidential Election were held today, who would you vote for?
   - Kamala Harris (Democratic Party)
   - Donald Trump (Republican Party)
   - Undecided
   - Prefer not to say

4. **If you answered "Undecided" in the previous question, who are you leaning towards?**
   - Kamala Harris (Democratic Party)
   - Donald Trump (Republican Party)
   - Prefer not to say

---

**Section 2: Key Issue Priorities**

5. What are the three most important issues influencing your vote in this election? (Select up to 3)
   - Economy
   - Immigration
   - Healthcare
   - National Security
   - Education
   - Climate Change

6. **How has the state of the U.S. economy influenced your voting choice?
   - Not at all
   - Somewhat
   - Greatly
   - Unsure

---

**Section 3: Voting History**

7. Did you vote in the 2020 U.S. Presidential Election?
   - Yes
   - No
   - Prefer not to say

8. If you voted in 2020, for whom did you vote?
   - Joe Biden (Democratic Party)
   - Donald Trump (Republican Party)
   - Prefer not to say

---

**Section 4: Voter Engagement**

9. How motivated are you to vote in the 2024 election compared to past elections?
   - Less motivated
   - About the same
   - More motivated
   - Unsure

10. Which of the following describes your voting method?
    - I plan to vote in person on Election Day
    - I plan to vote early in person
    - I plan to vote by mail
    - I have already voted (via early voting or mail-in)
    - I do not plan to vote

---

**Section 5: Demographic Questions**

11. What is your age range? (years)
    - 18-24
    - 25-34
    - 35-44
    - 45-54
    - 55-64
    - 65 or older
    - Prefer not to say

12. What gender do you identify as?
    - Male
    - Female
    - Non-binary
    - Prefer not to say

13. Which of the following best describes your race or ethnicity?
    - White or Caucasian
    - Black or African American
    - Hispanic or Latino of any race
    - Asian
    - Other or multiple races

14. What is the highest level of education you have completed?
    - Less than high school
    - High school diploma or GED
    - Some college
    - No degree
    - Associate degree
    - Bachelor’s degree
    - Graduate or professional degree
    - Prefer not to say

15. What is your current employment status?
    - Employed full-time
    - Employed part-time
    - Self-employed
    - Unemployed
    - Student
    - Retired
    - Prefer not to say

16. Do you consider yourself a supporter of any of the following political parties?
    - Democratic Party
    - Republican Party
    - Independent
    - Prefer not to say

---

**End Message**:

Thank you for completing this survey! Your responses are valuable in understanding voter opinions and trends in the lead-up to the 2024 U.S. Presidential Election.

\endgroup

# Code Styling {#sec-styling}

@@

# Reproducibility {#sec-reproducibility}

@@

# Acknowledgments {#sec-acknowledgement}
